The framing of active school travel in Ontario, Canada, as a health and built environment issue

### 2022-01-28 AST-Framing-Ontario/AST-Framing-Ontario.tex
Antonio's edits of latest version of paper

\usepackage{lineno} % add
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

@@

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\bibliographystyle{elsarticle-harv}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={The framing of active school travel in Ontario, Canada as a health and built environment issue},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

@@

% Pandoc citation processing

% Pandoc header
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\begin{document}
\begin{frontmatter}

  \title{The framing of active school travel in Ontario, Canada as a health and
built environment issue}
    \author[Some Department]{Author 1\corref{1}}
   \ead{author1@example.com} 
    \author[Some Department]{Author 2}
   \ead{author2@example.com} 
    \author[Another University]{Author 3\corref{2}}
   \ead{author3@example.com} 
    \author[Some Institute]{Author 4\corref{2}}
   \ead{author4@example.com} 
    \author[Some University]{Author 5}
   \ead{author5@example.com} 
    \author[Some Department]{Author 6}
   \ead{author6@example.com} 
      \address[Some Department]{Department, Street, City, Province, Postal Code}
    \address[Another University]{Department, Street, City, Province, Postal Code}
    \address[Some Institute]{Street, City, Province, Postal Code}
    \address[Some University]{Department, Street, City, Province, Postal Code}
      \cortext[1]{Corresponding Author}
    \cortext[2]{Equal contribution}

@@

Rates of walking and bicycling to school, commonly known as active
school travel (AST), have been declining in Canada and the United States
for decades (Rothman et al., 2018), with levels much lower than other
developed countries like The Netherlands (van Goeverden and de Boer,
2013), Denmark (Jensen, 2008), and Japan (Waygood and Taniguchi, 2020).
However, the appetite for AST may be high in many Canadian communities.
For instance, 40\% of children in a study conducted in Toronto, Ontario
who are driven to school would like to travel by bicycle instead
(Larouche et al., 2016). But the study reported that less than 3\%
actually do, despite the vast majority having access to a bicycle and
living within a short and bikeable distance to school. Another study
based in London, Ontario reported similar findings with respect to
children's preference for active travel (Larsen et al., 2012). From a
public health perspective, this represents a huge opportunity to improve
health and wellbeing if more children could use active modes to school.

In response to this trend, the Government of Ontario, in Canada, created
a fund in 2017 to support communities across the province in developing
AST initiatives. Green Communities Canada runs the program, called
\emph{Ontario Active School Travel}, and determines the projects that
will be funded with their support (Canada, 2020a). As of December 2021,
OAST has awarded over 2 million Canadian dollars and provided resources
to over 25 projects across Ontario. Many communities implemented school
travel planning (STP), along with encouragement activities like walking
school buses or developed resources for schools and parents, among other
actions (see Canada, 2020b). In this context, STP is a popular
``school-specific'' intervention led by a facilitator that brings
together a committee of stakeholders from diverse sectors including
education, planning, transportation, and public health to develop action
plans (Buliung et al., 2011; Mammen, M. R. Stone, Ron Buliung, et al.,
2014). The five-step process involves identifying barriers to AST based
on the local context and implementing approaches or activities following
the 5 E's that alleviate concerns about AST and make AST safer and more
convenient (Buliung et al., 2011; Lang et al., 2011).

How AST is framed by STP stakeholders to a target audience, like parents
or the general public, may raise awareness about the issue and influence
how walking and bicycling to school are perceived. Gamson and Modigliani
(1987) define a frame as a ``central organizing idea or story line that
provides meaning'' to a particular phenomenon. A frame can enable
individuals ``to locate, perceive, identify, and label'' information
pertaining to various dimensions of an issue (Goffman, 1974). Therefore,
stakeholder groups involved in STP efforts can play a role in shaping
public perception about AST in such a way that it attracts greater
attention and becomes a ``problem'' that needs to be addressed through
behaviour change or by new policies. We also contend that the ways in
which the AST issue is framed can affect the policy process, as outlined
in Kingdon's (1984) Multiple Streams Framework (MSF), through all three
streams (i.e., problem, policy, and politics). This framing can
ultimately impact financial contributions and potential solutions to
address the decline of AST in Canada.

After significant investment of human and financial resources to boost
rates of AST in Ontario over the past few years, we ask the following
questions: How is AST framed? What benefits of AST are communicated to
the public? What solutions are proposed to increase rates of AST? Do
proposed solutions focus primarily on the intrapersonal and
interpersonal factors, meaning that behaviour change must ultimately
result from the individual or household making different travel
decisions? Or is AST framed as an issue that must be addressed through
changes in the built environment or at the policy level?

In this paper, we use natural language processing to examine how three
STP stakeholder groups in Ontario (i.e., municipalities, school boards,
and transportation consortia) frame the issue of AST. We assembled a
corpus of texts from webpages that can be considered an important source
of information for the general public or parents who are interested in
AST. We examine word frequency, bigrams, and concordances in these
selected documents, and also identify key topics presented by each
stakeholder group. We then compare the findings from these documents to
a selection of studies on AST and explore the extent to which there is
concordance between the literature on AST and materials shared with the
public.

@@

The benefits of AST are valuable information to communicate to the
public in order to convey the importance of this issue. The desire to
increase AST in Canada is certainly warranted - there is compelling
evidence that children who actively commute to school earn physical and
mental health benefits. Faulkner et al.~(2009) concluded from their
systematic review that children who travel on foot or by bicycle to
school generally have higher levels of physical activity than their
peers who are driven to school. However, Schoeppe et al.~(2015) reported
no association between AST and physical activity. This relationship
could be dose-dependent, meaning that children would have to travel
longer distances to accumulate physical activity. A walking distance of
1000-1600 metres to school has been found to contribute to overall
levels of physical activity for boys (Faulkner et al., 2013). The daily
routine of travelling to school can be a good opportunity for children
to regularly build physical activity into their schedule (Mitra, 2013).
Research has also shown that using active modes to school contributes to
improvements in cardiovascular fitness (Børrestad et al., 2012).

More recently, there has been literature exploring the link between
transport and children's wellbeing (Waygood et al., 2020), with relevant
applications to the study of travel satisfaction (van den Berg et al.,
2020; Westman, Olsson, et al., 2017). Being driven to school reduces
community interactions for children which may impact their social
wellbeing (Waygood and Friman, 2015), whereas walking and bicycling give
children opportunities to socialize with friends or siblings (Michail et
al., 2021). This is something that children highly value about travel
(Zwerts et al., 2010) and indicates that social connections through
travel are important for their wellbeing. AST also provides
opportunities for children to engage with the natural environment (Fusco
et al., 2012; Romero, 2015) which can improve their mental or emotional
wellbeing.

@@

Factors that influence AST have been presented and organized using a
socio-ecological model (SEM) (Mitra, 2013) or systems model (Badland et
al., 2016) whereby children's travel behaviour is understood within the
context of their household, social, neighbourhood, and policy
environments. Individual characteristics of the child are also
determinants. The SEM comes from the field of public health and is a
useful framework for understanding complex health behaviours, including
travel habits such as walking or bicycling to school, because it
identifies multiple determinants that need to be addressed by
interventions to facilitate behaviour change. The consensus in the
literature is that various factors are interrelated and that
interventions should target multiple levels in order to increase levels
of healthy and physically active modes of travel to school (Mitra,
2013). Rather than describing an extensive list of factors that
influence AST and mode choice in Canada (e.g., Mammen et al., 2012;
Mitra, 2013; Rothman et al., 2018; Wilson et al., 2018), we discuss a
few potentially modifiable determinants that may be targeted for change
through STP.

At the individual level, older child age is often associated with AST
(Mammen et al., 2012; Stark et al., 2018; Wilson et al., 2018) as
parents typically escort younger children. There is evidence that gender
is a determinant of AST, with boys being more likely to travel using
active modes than girls, although this is not a strong or consistent
finding (Rothman et al., 2018; Schoeppe et al., 2015). Children's mode
choice to school is strongly influenced by their parents' travel
behaviours and the complexity of their household's travel needs (Buliung
et al., 2021), which indicates that shifting parental perceptions and
habits is important. Convenience and inclement weather have been cited
by parents as barriers to AST (Buliung et al., 2011). Parental
perceptions of the built or school environment (De Meester et al., 2014;
Panter et al., 2010) and their children's skills (Mammen et al., 2012)
also influence whether they allow their children to walk or bicycle to
school.

Distance between home and school is most strongly associated with AST
(Ikeda et al., 2018; Mammen et al., 2012; Pont et al., 2009; Rothman et
al., 2018) with less AST reported among children who have to travel
farther to school. Many studies have also found that the quality of the
built environment along the route to school and around the school site
(Ikeda et al., 2018; Rothman et al., 2021) and provision of active
travel infrastructure (Chen et al., 2018; Pont et al., 2009) facilitate
AST. Canadian youth report that they feel most safe bicycling on streets
in their neighbourhood or that have low volumes of traffic
({\textbf{???}}). Finally, concerns about traffic and strangers have
been reported by parents who drive their children to school (Mammen et
al., 2012), which highlights that the volume and speed of cars can be a
concern or deterrent for AST.

@@

School travel planning (STP) has been implemented in Canada since at
least the late 2000s. Within the STP process, facilitators generally
establish multi-sector committees who intervene at the participating
school through a range of activities related to the 5E's such as
\emph{education} strategies, \emph{encouragement} through in-person
events or programs, \emph{engineering} improvements to or around the
school site, and \emph{enforcement} of traffic speeds around schools
(Lang et al., 2011; Mammen, M. R. Stone, G. Faulkner, et al., 2014). The
first large-scale evaluation of STP as an intervention in Canada took
place at twelve schools across the country, including four in Ontario,
using parental surveys to measure changes in travel behaviour and
perceptions (Buliung et al., 2011). Since then, there have been other
assessments of both the efficacy of STP (Buttazzoni et al., 2019;
Mammen, M. R. Stone, G. Faulkner, et al., 2014) and the process of
implementing such programs in Ontario (Buttazzoni et al., 2018; Mammen
et al., 2015). STP facilitators have recommended that additional time
and resources are needed to improve the efficacy of STP (Mammen et al.,
2015), which highlights that long-term and sustained efforts driven at
the policy level are required to address declining rates of AST.

The ways in which AST is framed seem particularly important to shift
parental attitudes and perceptions given their reported influence on
children's travel mode to school. STP activities heavily focus on
education or encouragement (Buliung et al., 2011; Buttazzoni et al.,
2018; Mammen, M. R. Stone, G. Faulkner, et al., 2014), but parents may
not always be receptive to the goals of STP and may be resistant to
behaviour change (see Buttazzoni et al., 2018). Parents have been found
to express different understandings, language, and perceptions than
planners of the built environment determinants that influence school
travel (Buliung et al., 2021). This is also true when it comes to other
factors like convenience of different modes to school (Lang et al.,
2011). STP stakeholder groups must pay special attention to parents'
understanding of the decline of AST as a problem, which may affect their
receptivity to proposed solutions.

The ``central organizing idea or story line'' of AST, to apply the
definition from Gamson and Modigliani (1987), could also affect broader
support in the community. Municipal representatives are perceived to be
instrumental but the involvement of other stakeholder groups (e.g.,
busing consortium representatives and local residents) can be lacking
(Buttazzoni et al., 2018). In Ontario, it would be reasonable to say
that AST has become a policy issue on the education and public health
agendas over this time as evidenced by the financial contribution from
the provincial government. The support from a range of municipal
representatives (see Buttazzoni et al., 2018; Mammen et al., 2015)
demonstrates that the ``policy stream'' (see Kingdon, 1984) has been
well engaged. However, it is unknown to what degree the general public
(i.e., local residents) has been exposed to messaging about AST, or how
they view the AST issue, which could affect their participation in the
STP process as desired by STP facilitators (Buttazzoni et al., 2018).

The success of STP interventions would likely depend on parental
judgments of factors that are related to school travel, as well as
support from key policy makers. For example, parents have been found to
view mixed land use as conducive for driving, despite transport planners
viewing neighbourhoods with mixed uses as key for encouraging more
active travel (Buliung et al., 2021). Therefore, stakeholder groups
involved in STP must make important choices about the proposed solutions
and potential benefits of AST that ought to be communicated to parents
and the general public to convey its importance as a policy issue and to
facilitate adoption of AST. Publicly available content about AST,
therefore, needs to effectively engage multiple audiences on this policy
issue including parents, children, politicians, and school
representatives. This information should reflect current knowledge from
research on school travel, plus content specific to local factors that
influence AST, so that the challenges of AST are adequately defined and
the opportunities or solutions to address the problems are clear.
However, it is unknown to what degree STP materials presented to the
general public reflect the scope of evidence from the literature on AST.

@@

\caption{\label{tab:policy-documents}\label{tab:search-results}Search results from the main STP stakeholder groups.}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{>{}l|l|>{}l}
\toprule
Stakeholder & Total & Included\\
\midrule
\cellcolor{gray!6}{\textbf{School boards}} & \cellcolor{gray!6}{62} & \cellcolor{gray!6}{32}\\
\textbf{Municipalities} & 62 & 28\\
\cellcolor{gray!6}{\textbf{Transportation consortia}} & \cellcolor{gray!6}{39} & \cellcolor{gray!6}{9}\\
\bottomrule
\end{tabular}}
\end{table}

@@

\textbf{To be completed by Dr.~Paez}

@@

A multi-step process was conducted to ensure that the analysis captured
as much text as possible from both the policy documents (n = 64) and
academic papers (n = 233). To begin, the webpages, which were manually
downloaded in portable document format (PDF), were trimmed so that pages
that only consisted of tables, figures, or references were removed. Many
academic papers were in a two-column format, which is not ideal for
conversion to \texttt{txt}. We adapted a procedure
(https://stackoverflow.com/questions/42541849/extract-text-from-two-column-pdf-with-r)
to read the two-column PDF documents so that they would be converted
correctly. Four academic papers did not join sufficiently and were taken
out of the corpus due to the substantial time required to manually
correct their inconsistencies.

Next, we converted the trimmed PDF documents into \texttt{txt} files so
that they could be imported in R for analysis. We then proceeded to a
manual cleaning phase where we removed any remaining tables, figures,
references, headers/footings, and captions that could not be trimmed.
Manual corrections were also required for certain pages in academic
papers that remained in two-column format after the conversion process.
This typically occurred on pages that had a table or figure which
disrupted the text. Finally, we reviewed all of the documents to remove
hyphenation by line breaks and to keep hyphenated words together on the
same line. Any ligatures (e.g., combinations of characters or letters
that were not properly detected during the conversion process) were
fixed by replacing the unicode sequence of character by inserting the
missing sequence of characters.

We also manually removed any extraneous material in the academic papers
that did not pertain to AST specifically. This included footnotes,
references, acknowledgments, and conflict of interest statements in the
academic papers. We removed all phone numbers, inserted links to other
webpages, personal names, and content not to specific to AST from the
policy documents that were retrieved from the websites of school boards,
municipalities, and transportation consortia.

In the final step, we removed all blank spaces, punctuation,
capitalization, and numbers. English stop words, which are common words
such as \emph{and} or \emph{the} as identified in a predetermined list
by Lewis et al.~({\textbf{???}}) and other frequent terms in the
documents like ``school'' and specific location names, were removed from
the corpora.

@@

This paper is an example of open and reproducible research that uses
only open software. All data were obtained from publicly available
sources and organized in the form of a data package. Following best
practices in spatial data science ({\textbf{???}}), the code and data
needed to reproduce or conduct a similar analysis for other regions in
North America or elsewhere are available for download.

@@

@@

\caption{\label{fig:city-visual}Most common bigrams found in the municipal or regional government documents.}\label{fig:city-visual}
\end{figure}

@@

\begin{quote}
\emph{Walking to school is a great way to add physical activity into
your child's busy day.} (Region of Haldimand-Norfolk Region)
\end{quote}

@@

\hypertarget{framing-in-stp-documents}{%
\subsection{6.1. Framing in STP
documents}\label{framing-in-stp-documents}}

>>>>

\usepackage{lineno} % add

@@

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\bibliographystyle{elsarticle-harv}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={The framing of active school travel in Ontario, Canada, as a health and built environment issue},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

@@

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Pandoc citation processing
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
% for Pandoc 2.8 to 2.10.1
\newenvironment{cslreferences}%
  {}%
  {\par}
% For Pandoc 2.11+
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\begin{document}

\begin{frontmatter}

  \title{The framing of active school travel in Ontario, Canada, as a
health and built environment issue}
    \author[Some Department]{Author 1\corref{1}}
   \ead{author1@example.com} 
    \author[Some Department]{Author 2}
   \ead{author2@example.com} 
    \author[Another University]{Author 3\corref{2}}
   \ead{author3@example.com} 
    \author[Some Institute]{Author 4\corref{2}}
   \ead{author4@example.com} 
    \author[Some University]{Author 5}
   \ead{author5@example.com} 
    \author[Some Department]{Author 6}
   \ead{author6@example.com} 
      \address[Some Department]{Department, Street, City, Province,
Postal Code}
    \address[Another University]{Department, Street, City, Province,
Postal Code}
    \address[Some Institute]{Street, City, Province, Postal Code}
    \address[Some University]{Department, Street, City, Province, Postal
Code}
      \cortext[1]{Corresponding Author}
    \cortext[2]{Equal contribution}

@@

Rates of walking and bicycling to school, commonly known as active
school travel (AST), have been declining in Canada and the United States
for decades (Rothman et al., 2018). However, the appetite for AST may be
high in many Canadian communities. For instance, in a study conducted in
Toronto, Ontario, 40\% of children who are driven to school would like
to travel by bicycle instead (Larouche et al., 2016). The study also
reported that less than 3\% of respondents actually cycled, despite the
vast majority having access to a bicycle and living within a short and
bikeable distance to school. Another study based in London, Ontario,
reported similar findings with respect to children's preference for
active travel (Larsen et al., 2012). From a public health perspective,
increasing the use of active modes to school represents a major
opportunity to improve children's health and wellbeing.

The Government of Ontario, in Canada, has striven to promote this trend,
and created a fund in 2017 to support communities across the province to
develop AST initiatives. Green Communities Canada runs the program,
called \emph{Ontario Active School Travel} (Canada, 2020a). As of
December 2021, OAST has awarded over CAD 2 million and provided
resources to over 25 projects across Ontario. Many communities
implemented school travel planning (STP), along with encouraging
activities like walking school buses and developed resources for schools
and parents, among other actions (see Canada, 2020b). In this context,
STP is a popular ``school-specific'' intervention led by a facilitator
who brings together a committee of stakeholders from diverse sectors
including education, planning, transportation, and public health to
develop action plans (Buliung et al., 2011; Mammen et al., 2014a). The
five-step process involves identifying barriers to AST based on the
local context and implementing approaches or activities following the
4E's that make AST safer and more convenient: 1) setup of the program;
2) data collection and problem identification; 3) action planning; 4)
implementation; 5) monitoring and evaluation (Buliung et al., 2011; Lang
et al., 2011).

How AST is framed by STP stakeholders to a target audience, like parents
or the general public, may raise awareness of the issue and influence
how walking and bicycling to school are perceived. Gamson and Modigliani
(1987) define a frame as a ``central organizing idea or story line that
provides meaning'' to a particular phenomenon. A frame can enable
individuals ``to locate, perceive, identify, and label'' information
pertaining to various dimensions of an issue (Goffman, 1974). Therefore,
stakeholders involved in STP efforts can play a role in shaping public
perception about AST in such a way that it attracts greater attention
and is recognized as a ``problem'' that needs to be addressed through
behaviour change or new policies.

After significant investment of human and financial resources to boost
rates of AST in Ontario over the past few years, we ask this central
question: How is AST framed to the public? The following questions are
subsidiary and help us address this main question: What benefits of AST
are communicated to the public? What solutions are proposed to increase
rates of AST?

In this paper, we use natural language processing to examine how three
STP stakeholder groups in Ontario (municipalities, school boards, and
transportation consortia) frame the issue of AST. We assembled a corpus
of texts from public sources that present information for the general
public or parents who are interested in AST. We examine word frequency,
bigrams, and concordances in these selected documents, and also identify
key topics presented by each stakeholder group. We then compare the
findings from these documents to a selection of studies on AST and
explore the extent to which there is concordance between the literature
on AST and materials shared with the public.

@@

There is value in communicating the benefits of AST to the public in
order to convey the importance of this issue. The desire to increase AST
in Canada is certainly warranted - there is compelling evidence of the
physical and mental health benefits that children who actively commute
to school attain. Faulkner et al. (2009) concluded from their systematic
review that children who travel on foot or by bicycle to school
generally have higher levels of physical activity than their peers who
are driven to school. However, Schoeppe et al. (2015) reported no
association between AST and physical activity in their cross-sectional
study. This relationship could be dose-dependent, meaning that children
would have to travel longer distances to accumulate physical activity. A
walking distance of 1000-1600 metres to school has been found to
contribute to overall levels of physical activity for boys (Faulkner et
al., 2013). The daily routine of travelling to school can be a good
opportunity for children to regularly build physical activity into their
schedule (Mitra, 2013). Research has also shown that using active modes
to school contributes to improve cardiovascular fitness (Børrestad et
al., 2012).

More recently, the literature has explored the link between transport
and children's wellbeing (Waygood et al., 2020), with relevant
applications to the study of travel satisfaction (van den Berg et al.,
2020; Westman et al., 2017b). Being driven to school reduces community
interactions for children, which may negatively impact their social
wellbeing (Waygood and Friman, 2015). In contrast, walking and cycling
give children opportunities to socialize with friends or siblings
(Michail et al., 2021). This is something that children seem to value
(Zwerts et al., 2010) and that tracks with findings among university
students (Paez and Whalen, 2010). More generally, social connections
through travel appear to be important for wellbeing. Furthermore, AST
can provide opportunities for children to engage with natural
environments (Fusco et al., 2012; Romero, 2015).

@@

Factors that influence AST have been presented and organized using a
socio-ecological model (SEM) (Mitra, 2013) or systems model (Badland et
al., 2016) whereby children's travel behaviour is understood within the
context of household, social, neighbourhood, and policy environments.
The SEM comes from the field of public health and is a useful framework
for understanding complex health behaviours, such as walking and cycling
habits, This model helps to identify multiple determinants that need to
be addressed by interventions to facilitate behaviour change. The
consensus in the literature is that interventions should target multiple
levels in order to increase levels of healthy and physically active
modes of travel to school (Mitra, 2013). Rather than describing an
extensive list of factors that influence AST and mode choice in Canada
(e.g., Mammen et al., 2012; Mitra, 2013; Rothman et al., 2018; Wilson et
al., 2018), we discuss a few potentially modifiable determinants that
may be targeted for change through STP.

At the individual level, age is often positively associated with AST
(Mammen et al., 2012; Stark et al., 2018; Wilson et al., 2018). There is
evidence that gender is a determinant of AST, and that boys are more
likely to travel using active modes than girls, although this is not a
strong or consistent finding (Rothman et al., 2018; Schoeppe et al.,
2015). Children's mode choice to school is strongly influenced by their
parents' travel behaviours and the complexity of their household's
travel needs (Buliung et al., 2021), as well as the frequency of
parental support (\textbf{mahDoesParentalSupport2017a?}). This indicates
that shifting parental perceptions and habits is important. Convenience
and inclement weather have been cited by parents as barriers to AST
(Buliung et al., 2011). Parental perceptions of the built or school
environment (De Meester et al., 2014; Panter et al., 2010) and their
children's skills (Mammen et al., 2012;
\textbf{faulknerWhatQuickestEasiest2010?}) also influence mode choice to
school.

Distance between home and school is most strongly associated with AST
(Ikeda et al., 2018; Mammen et al., 2012; Pont et al., 2009; Rothman et
al., 2018) with less AST reported among children who have to travel
farther to school - low density suburban development tends to increase
the need for movement, and the reliance on motorized modes (Farber and
Páez, 2011). Many studies have also found that the quality of the built
environment along the route to school and around the school site (Ikeda
et al., 2018; Rothman et al., 2021) and provision of active travel
infrastructure (Chen et al., 2018; Pont et al., 2009) facilitate AST.
Canadian youth report that they feel most safe bicycling on streets in
their neighbourhood or that have low volumes of traffic
(\textbf{TACbikeinfra2020?}). Finally, concerns about traffic and
strangers have been reported by parents who drive their children to
school (Mammen et al., 2012), which highlights that the volume and speed
of cars can be a concern or deterrent for AST.

@@

School travel planning (STP) has been implemented in Canada since at
least the late 2000s. Within the STP process, facilitators generally
establish multi-sector committees who intervene at the participating
school through a range of activities related to the 4E's consisting of
\emph{education} strategies, \emph{encouragement} through in-person
events or programs, \emph{engineering} improvements to or around the
school site, and \emph{enforcement} of traffic speeds around schools
(Lang et al., 2011; Mammen et al., 2014b). The first large-scale
evaluation of STP as an intervention in Canada took place at twelve
schools across the country, including four in Ontario, using parental
surveys to measure changes in travel behaviour and perceptions (Buliung
et al., 2011). Assessments of the efficacy of STP (Buttazzoni et al.,
2019; Mammen et al., 2014b) indicate that it has the potential to
encourage behaviour change and adoption of AST. STP facilitators have
recommended that additional time and resources are needed to improve the
efficacy of STP (Mammen et al., 2015), which highlights that long-term
and sustained efforts driven at the policy level are required to address
declining rates of AST.

How AST is framed seems particularly important to shift parental
attitudes and perceptions given their reported influence on children's
travel mode to school. STP activities heavily focus on education or
encouragement (Buliung et al., 2011; Buttazzoni et al., 2018; Mammen et
al., 2014b), but parents may not always be receptive to the goals of STP
and may be resistant to behaviour change (Buttazzoni et al., 2018).
Parents have been found to express different understandings, language,
and perceptions than planners of how the built environment can influence
school travel (Buliung et al., 2021). This is also true when it comes to
other factors like convenience of different modes to school (Lang et
al., 2011). STP stakeholders must pay special attention to parents'
understanding of the decline of AST as a problem, which may affect their
receptivity to proposed solutions.

The ``central organizing idea or story line'' of AST, to apply the
definition of framing from Gamson and Modigliani (1987), could also
affect broader support in the community. Municipal representatives are
perceived to be instrumental but the involvement of other stakeholder
groups (e.g., busing consortia representatives and local residents) can
be lacking (Buttazzoni et al., 2018). In Ontario, it is reasonable to
say that AST has become a policy issue on the education and public
health agendas, supported, albeit in a modest way, by financial
contributions from the provincial government\footnote{To put the CAD 2
  million invested by the province in the Ontario Active School Travel
  program, a city such as Hamilton budgeted in excess of CAD 91 million
  on road maintenance and construction in 2016 alone; see:
  https://www.hamilton.ca/budget-finance/city-budgets/2016-tax-and-rate-budgets}.
Still, the support from a range of municipal representatives (see
Buttazzoni et al., 2018; Mammen et al., 2015) demonstrates that the AST
issue is on the political agenda. However, it is unknown to what degree
the general public (i.e., local residents) has been exposed to messaging
about AST.

The success of STP interventions would likely depend on parental
judgments of factors that are related to school travel, as well as
support from key policy makers. For example, parents have been found to
view mixed land use as conducive for driving, despite transport planners
viewing neighbourhoods with mixed uses as key for encouraging more
active travel (Buliung et al., 2021). Therefore, stakeholders involved
in STP must make conscious choices about the proposed solutions and
potential benefits of AST that are communicated to parents and the
general public to convey its importance as a policy issue and to
facilitate adoption of AST. Publicly available content about AST needs
to effectively engage multiple audiences on this policy issue, including
parents, children, politicians, and school representatives. This
information should reflect current knowledge from research on school
travel, plus content specific to local factors that influence AST, so
that the challenges of AST are adequately defined and the opportunities
or solutions to address the problems are clear.

@@

\caption{\label{tab:policy-documents}\label{tab:search-results}Search results from the main STP stakeholder groups.}
\centering
\begin{tabular}[t]{>{}l|l|>{}l}
\toprule
Stakeholder & Total & Retrieved\\
\midrule
\cellcolor{gray!6}{\textbf{School boards}} & \cellcolor{gray!6}{62} & \cellcolor{gray!6}{32}\\
\textbf{Municipalities} & 62 & 28\\
\cellcolor{gray!6}{\textbf{Transportation consortia}} & \cellcolor{gray!6}{39} & \cellcolor{gray!6}{9}\\
\bottomrule
\end{tabular}
\end{table}

@@

We conducted a search on Web of Science for scholarly papers on the
topic of travel to school. We limited our search to the fields of public
health, transportation, planning, urban studies, and geography. This
initial corpus was curated by the authors to ensure that all documents
were relevant; for example, some hits in our search were not full papers
but conference reports, or they dealt with travel to school only
peripherally. After this step, we assembled a collection of 227 journal
articles for analysis.

@@

A multi-step process was conducted to ensure that the analysis captured
as much text as possible from both the policy documents (n = 64) and
academic papers (n = 227). To begin, the webpages, which were manually
downloaded in portable document format (PDF), were trimmed so that pages
that only consisted of tables, figures, or references were removed. Many
academic papers were in a two-column format, which is not ideal for
conversion to \texttt{txt}. Two-column PDF documents were processed to
extract the text in the correct sequence. Four academic papers did not
join sufficiently and were taken out of the corpus due to the
substantial time required to manually correct their inconsistencies.

Next, we converted the trimmed PDF documents into \texttt{txt} files so
that they could be imported for analysis in \texttt{R}. The next phase
was to manually remove any remaining tables, figures, references,
headers/footings, and captions that could not be trimmed. Manual
corrections were also required for certain pages in academic papers that
remained in two-column format after the conversion process. This
typically occurred on pages that had a table or figure that disrupted
the text. Finally, we reviewed all of the documents to remove
hyphenation by line breaks and to keep hyphenated words together on the
same line. Any ligatures (e.g., combinations of characters or letters
that were not properly detected during the conversion process) were
fixed.

We also manually removed any extraneous material in the academic papers
that did not pertain to AST specifically. This included footnotes,
references, acknowledgments, and conflict of interest statements in the
academic papers. We removed all phone numbers, inserted links to other
webpages, personal names, and content not specific to AST from the
policy documents that were retrieved from the websites of school boards,
municipalities, and transportation consortia.

In the final step, we removed all blank spaces, punctuation,
capitalization, and numbers. English stop words, which are common words
such as \emph{and} or \emph{the} as identified in a predetermined list
by Lewis et al. (Feinerer and Hornik, 2020) and other frequent terms in
the documents like ``school'' and specific location names, were removed
from the corpora.

@@

This paper is an example of open and reproducible research that uses
only open software. All data are organized in the form of an open data
product (Arribas-Bel et al., 2021). Following best practices in spatial
data science (\textbf{brunsdon2020opening?}), the code and data needed
to reproduce our research or conduct a similar analysis for other
regions are available for download.

@@

Examination of document frequency reveals terms that are not present in
all policy documents. This suggests that although documents pertain to
the subject of school travel, not all stakeholders across Ontario are
disseminating information about AST. For example, a document may discuss
active travel but not to school. Some documents present school travel
options by bus but not by active modes. We manually searched the policy
corpus and found that 48\% of documents mention AST and 16\% mention
STP. This confirms that many municipalities, school boards, and
transportation consortia do not promote AST through their webpages or
indicate their involvement in the STP process. Instead, inclement
weather and its impacts on busing is a common topic addressed in school
board and transportation consortia documents.

@@

\caption{\label{fig:city-visual}Most common bigrams found in the municipal and regional government documents.}\label{fig:city-visual}
\end{figure}

@@

@@

\hypertarget{current-ast-frames-in-ontario}{%
\subsection{6.1. Current AST frames in
Ontario}\label{current-ast-frames-in-ontario}}
